---
title: "Run differential abundance analysis between elite participants and matched others"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction

Run differential abundance analysis between elite and matched others.

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 10))

# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(DESeq2)
library(reshape2)
```

## Import data
```{r}
d.in <- "~/IKMB/projects/sport/data.clean"

meta <- "elite.group.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read_tsv()
micro <-  "micro.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read.delim(stringsAsFactors = F)
micro.r <- "micro.r.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read.delim(stringsAsFactors = F)
tax <- "tax.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read.delim(stringsAsFactors = F)
pca <- "~/IKMB/projects/sport/analysis/25.to.publish/6.diet/pca.rds" %>% 
  readRDS()

```
#Meta
```{r}
meta <- meta %>% 
  inner_join(pca, by = "Seq.ID")
```

# Filter data for group and period

## Filter microbiome

```{r}
micro <- micro %>%
  rownames_to_column("rownames") %>% # match microbiome data with meta
  left_join(data.frame(rownames = meta$Seq.ID),
            .,
            by = "rownames") %>%
  column_to_rownames("rownames")
micro <- micro[,colSums(micro)> 0]

micro.r <- micro.r %>%
  rownames_to_column("rownames") %>% # match microbiome data with meta
  left_join(data.frame(rownames = meta$Seq.ID),
            .,
            by = "rownames") %>%
  column_to_rownames("rownames")
micro.r <- micro.r[,colSums(micro.r)> 0]
```

# Perform DESEq2 routine


## Construct a *DESeqDataSet* object

```{r}
micro.t <- t(micro) %>% 
  data.frame()

dds <- DESeqDataSetFromMatrix(countData = micro.t,
                                colData = meta,
                                design = ~ PC1 + Group)
```

## Gather bacteria to keep
I will only keep bacteria that are present in at least 10% of samples and mean base of 10.
```{r}
#Set cutoff core
cut.off.core <- 0.10 #proportion
min.samples <- nrow(micro) * cut.off.core
print(min.samples)
# Get ASVs that pass the core cutoff
core <- micro > 0
core <- core %>% colSums()
core <- core[core > min.samples]
core <- names(core)
#Set cut-off base mean
mean.cut.off <- 10
# Set mean abundance
mean.ab <- micro.r %>% colSums()
mean.ab <- mean.ab / nrow(micro.r)
mean.ab <- mean.ab[mean.ab > mean.cut.off]
mean.ab <- names(mean.ab)
asv.to.keep <- intersect(core, mean.ab)
print(length(asv.to.keep))
```

# Estimate size factor
```{r}
dds <- estimateSizeFactors(dds, "poscounts") # necessary due to error with  geometric mean equal to zero
idx <- rownames(micro.t) %in% asv.to.keep
names(idx) <- rownames(micro.t)
dds <- dds[idx,]
```

## Do the test

```{r, cache = F}
dds <- DESeq(dds)
```
### View

```{r}
plotMA(results(dds) )
```

## Check results
```{r}
res <- results(dds) %>%
   data.frame()

ggplot(res, aes(x = pvalue)) +
  geom_histogram(bins = 50) +
  ggtitle("DESeq2 p-value distribution")

```

## Include info

```{r}
res <- res %>%
  rownames_to_column("ASV") %>%
  arrange(padj) %>% 
  left_join(tax %>% select(ASV, Phylum, Genus), by = "ASV") 
```

### View
```{r}
head(res, 50)
```

## Plot log fold change

```{r}
res %>%
  filter(padj < 0.05) %>% 
  ggplot(aes(x = Genus, y = log2FoldChange, colour = Phylum)) +
  geom_point(size =3, alpha =0.8) +
  ggpubr::rotate_x_text() +
  coord_flip() +
  theme(legend.position = "bottom") +
  labs(y= "log2 fold change\n(Control/Elite)", x = "")
# Save plotting info
paste0(d.out, "/", "to.plot.diff.rds") %>%
  saveRDS(res, .)
```


# Save

```{r}
paste0(d.out, "/","dseq.res.stats.tsv") %>%
  write_tsv(res, .)
```

# Session information

```{r}
sessionInfo()
```
