---
title: "Run comparison between microbiomes' diversity of elite participants and matched others"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction
Compare microbiome diversity between elite and matched others

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(vegan)
library(reshape2)
library(broom) # plotting residuas vs fit nicely
library(spaa) #convert dist to list
library(stats)
```

## Import data
```{r}
d.in <- "~/IKMB/projects/sport/data.clean"

meta <- "elite.group.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read_tsv()
micro <-  "micro.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read.delim(stringsAsFactors = F)
micro.r <- "micro.r.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read.delim(stringsAsFactors = F)
alpha <- "alpha.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read_tsv()
pca <- "~/IKMB/projects/sport/analysis/25.to.publish/6.diet/pca.rds" %>% 
  readRDS()

```
#Meta
```{r}
meta <- meta %>% 
  inner_join(pca, by = "Seq.ID")
```

# Filter data for group and period

## Filter microbiome

```{r}
micro <- micro %>%
  rownames_to_column("rownames") %>% # match microbiome data with meta
  left_join(data.frame(rownames = meta$Seq.ID),
            .,
            by = "rownames") %>%
  column_to_rownames("rownames")
micro <- micro[,colSums(micro)> 0]

micro.r <- micro.r %>%
  rownames_to_column("rownames") %>% # match microbiome data with meta
  left_join(data.frame(rownames = meta$Seq.ID),
            .,
            by = "rownames") %>%
  column_to_rownames("rownames")
micro.r <- micro.r[,colSums(micro.r)> 0]
```

## Filter alpha diversity

```{r}
alpha <- alpha %>%
  right_join(., meta, by = "Seq.ID")
```


# Alpha diversity


```{r}
alpha <- alpha %>%
  right_join(., meta)
```

## Select alpha tests

```{r}
alpha.indices <- c("Chao1", "InvSimp")
```

## Visualize


```{r}
alpha %>%
  select(!!alpha.indices, Group) %>%
  melt() %>%
  dplyr::rename("alpha.div" = "variable") %>%
  ggplot(., aes(x = Group,
                     y = value,
                     color = Group)) +
  geom_boxplot(coef = 6) +
  geom_jitter() +
  theme(legend.position = "bottom") +
  facet_wrap(~ alpha.div, scales = "free") +
  labs(y = "Index")
#Save plotting info
paste0(d.out, "/", "to.plot.alpha.rds") %>%
  saveRDS(alpha, .)
```


### Prepare function to test


```{r}
do.lm.test <- function(data, variable){
    df <- data %>%
    # Arrange, so the pairs can be recognized by the test
      arrange(ParticipantID) %>%
      arrange(Group)
    # Perform test
    #include ParticipantID in the formula because I have people that contributed with more samples than others
    f <- as.formula(paste("log(", variable,")", "~ PC1 + Group"))
    model <- lm(f, data = df) 
    sig <- anova(model) %>% 
      tidy() %>%
    mutate(index = variable,
           anova = "type I")
    # Calculate get coefficient
    model.coef <- model %>% 
      coef %>%
      data.frame(coef = .) %>% 
      rownames_to_column() %>%
      filter(rowname == "GroupElite") %>% 
      rename("coef.term" = "rowname")
      # Calculate get coefficient confint
    model.coef.confi <- model %>%
      confint() %>%
      as.data.frame() %>% 
      magrittr::set_colnames(c("conf.low", "conf.high")) %>% 
      rownames_to_column() %>%
      filter(rowname == "GroupElite") %>% 
      select(-rowname)
    

  # Gather results
      r <- sig %>%
        filter(term == "Group") %>% 
        bind_cols(model.coef, model.coef.confi)

      return(r)
}

```


### Perform tests
```{r}
alpha.stats <- NULL
for (i in alpha.indices){
  alpha.stats <- bind_rows(alpha.stats,
                             do.lm.test(alpha,
                                       i))
}
```

### View tests

```{r}
alpha.stats
```

# Beta diversity


The aim is to see if there is any difference in beta diversity due to time.

##Making the plot

```{r}
mds <- micro.r %>%
  vegdist('bray') %>%
  metaMDS(.,autotransform = F)
```

## Visualize the NMDS plot


```{r, fig.height=12, fig.width=12}

to.plot.beta <- mds$points %>%
  data.frame() %>%
  rownames_to_column("Seq.ID") %>%
  inner_join(alpha) %>%
  mutate(ParticipantID = factor(ParticipantID))
to.plot.beta %>% 
  #plot
  ggplot(aes(x= MDS1, y = MDS2, colour =Group, group = ParticipantID)) +
  geom_point(aes(colour = Group), size = 3) +
  # add arrow
  geom_line(alpha =0.4, size = 2.5) +
  #legend nicer, save space
  theme(legend.position="bottom")
#Save plotting info
paste0(d.out, "/", "to.plot.beta.rds") %>%
  saveRDS(to.plot.beta, .)
```

## Stats

### Get distances

```{r}
dis <- micro.r %>%
  vegdist('bray') 
```
### Perform stats

```{r}
beta.stats <-  adonis2(dis ~ PC1 + Group , data = alpha, by = "terms") %>% 
  tidy() %>% 
  mutate(stat.type = "F",
         method = "adonis.2")
```

### Perform stats

```{r}
beta.disp <-  betadisper(dis, factor(alpha$Group), bias.adjust = T) %>% 
  permutest(pairwise = F, permutations = 999) %>% 
  .[["tab"]] %>% 
  data.frame() %>%
  rename("p.value" = "Pr..F.",
         "permutations" = "N.Perm",
         "statistic" = "F",
        "SumOfSqs" = "Sum.Sq",
        "df" = "Df") %>% 
  rownames_to_column("term") %>% 
  mutate(stat.type = "F",
         method = "betadisper.permutest")
```



### View tests

```{r}
beta.stats
beta.disp
```

## Save outputs

```{r}
paste0(d.out, "/", "alpha.stats.tsv") %>%
  write_tsv(alpha.stats, .)

paste0(d.out, "/", "beta.stats.tsv") %>%
  write_tsv(beta.stats, .)

paste0(d.out, "/","beta.disp.tsv") %>%
  write_tsv(beta.disp, .)


```


# Session information

```{r}
sessionInfo()
```
