---
title: "Visualize and test biometric variables"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction
I will visualize and test biometric variables.


## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 16))

# Set seed for reproducibility
set.seed(13)

# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}


# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(magrittr)
library(ggplot2)
library(tidyverse)
library(broom)
library(reshape2)
library(readODS)
```
# Import data

```{r}
meta <- "~/IKMB/projects/sport/data.clean/sampling.data.tsv" %>% 
  read_tsv() %>% 
  select(ParticipantID, Group) %>% 
  unique()


biometrics <- "~/IKMB/projects/sport/data.clean/biometrics.tsv" %>% 
  read_tsv()
```

# Add group information

```{r}
biometrics <- biometrics %>% 
  inner_join(meta %>%  select(ParticipantID, Group) %>% unique(), by = "ParticipantID")
```



# Test function

```{r}

do.w.test <- function(data, test, group){
  
  # Select column to test
  column <- "Group"
  # Select data
  df <- data %>% 
    rename("Grouping" = column) %>% 
    filter(Test == test,
           Grouping == group)
  
  #Remove NA
  df <- df %>%
    filter(!is.na(Value))
  
  # Select only paired Participants
  odd.participants <- count(df, ParticipantID) %>%
    filter(n != 2) %>% 
    pull(ParticipantID)
  df <- df %>% 
    filter(!ParticipantID %in% odd.participants)
  ## Arrange, so the pairs can be recognized by the test
  df <- df %>% 
    arrange(ParticipantID) %>%
    arrange(Timepoint)
  # Make the formula
  f <- as.formula("Value ~ Timepoint")
  
  # Perform wilcoxon
  wcx <- wilcox.test(f, data = df, paired = T) %>%
    tidy() %>%
    mutate(test = test,
           group = group,
           n = nrow(df)/2)
  
  # Calculate r (rank-biserial correlation), version 1 ( After Kerby 2014)
  #  r = f – u
  sum.ranks  <- df %>% 
    select(Value, Timepoint, ParticipantID) %>% 
    dcast(ParticipantID  ~ Timepoint, value.var = "Value") %>% 
    # get changes. ties are averaged (default)
    mutate(change = After - Before) %>% 
    # Get ranks of absolute values of the change score
    mutate(rank = rank(abs(change))) %>%
    #get direction
    mutate(direction = if_else(change > 0,
                               "favorable",
                               "unfavorable")) %>% 
    group_by(direction) %>%
    #calculate sum of favorable and unfavorable ranks
    summarise(sum.ranks = sum(rank)) %>%
    #Get proportions of fav and unfav
    ungroup() %>%
    mutate(sum.ranks.proportion = sum.ranks/sum(sum.ranks)) %>% 
    #Make sure sum of ranks has all the directions
    right_join(data.frame(direction = c("favorable",
                                        "unfavorable"),
                          strinsgAsFactors = F),
               by = "direction")
  #Conver NAś to zero in sum of ranks
  sum.ranks[is.na(sum.ranks)] <- 0
  
  rfu <- sum.ranks %>% 
    # get matched-pairs rank-biserial correlation as r = f – u
    summarize(r = filter(., direction == "favorable") %>%
                pull(sum.ranks.proportion) - filter(., direction == "unfavorable") %>% 
                pull(sum.ranks.proportion)) %>% 
    pull(r)
  
  # Gather results
  wcx <- wcx %>% 
    mutate(r = rfu,
           favorable = sum.ranks %>%
             filter( direction == "favorable") %>% 
             pull(sum.ranks.proportion),
           unfavorable = sum.ranks %>%
             filter( direction == "unfavorable") %>% 
             pull(sum.ranks.proportion)
    )
  
  return(wcx) 
}
```

# Set variables to test
```{r}
groups <- biometrics %>% pull(Group) %>% unique()
tests <- biometrics %>% pull(Test) %>% unique()
tests
```
# Remove NAs
```{r}
biometrics <- biometrics %>%
  filter(!is.na(Value))
```

# Correct 0 values to NA

```{r}

biometrics <- biometrics %>% 
  mutate(Value = if_else(Value == 0,
                         NA_real_,
                         Value))
```


# Perform the tests

```{r, warning = F}
res <- NULL
for (g in groups){
  for(t in tests){
    tryCatch({
      res <- bind_rows(res,
                       do.w.test(biometrics, t, g))},
      error = function(e){print(paste("error", g, t))})
  }
}
```

# Adjusting

```{r}
res <- res %>% 
  group_by(group) %>% 
  mutate(p.adj = p.adjust(p.value, "fdr")) %>% 
  select(test, group, p.adj, p.value, everything()) %>% 
  arrange(p.adj)
```

Note that, as in other tests, the adjustment of multiple testing was done withing each group.


# Do some arrangings on the results table

```{r}
res <- res %>% 
  ungroup() %>% 
  arrange(group, test) %>% 
  mutate(group = factor(group, levels= c("Control","Endurance" , "Strength" )))
```

# Save results in a table
```{r}
"statistical.results.tsv" %>% 
  paste0(d.out, "/",.) %>% 
  write_tsv(res,.)
```
# Save results in a table
```{r}
"to.publish.statistical.results.tsv" %>% 
  paste0(d.out, "/",.) %>% 
  write_tsv(res %>% select(-favorable, -unfavorable) %>% 
              filter(group %in% c("Control","Endurance" , "Strength")),.)
```


# Overview

## All results


```{r, fig.height=9}

res %>%
  filter(group %in% c("Control","Endurance" , "Strength")) %>% 
  select(test, group, p.adj, p.value, r) %>%
  droplevels() %>% 
  mutate(sig.after.adjusting = if_else(p.adj < 0.05,
                                       "yes",
                                       "no"),
         sig.after.adjusting = if_else(is.na(sig.after.adjusting),
                                       "no",
                                       sig.after.adjusting)) %>% 
  ggplot(aes(x = group, y = test, fill = r))+
  geom_tile() +
  scale_fill_gradient2(low="dodgerblue4", mid="white", high="firebrick4", 
                       midpoint=0) +
  ggpubr::rotate_x_text() +
  geom_point(aes(color = sig.after.adjusting), size =3) +
  guides(color=guide_legend(title="FDR < 0.05")) +
  labs( x = "Group", y = "Test")
```


# Detailed plots

## Significants

```{r}
label.df <- res %>%
  filter(p.adj < 0.05) %>% 
  rename("Test" = "test",
         "Group" = "group") %>% 
  mutate(label = "*") %>% 
  select(Test, Group, label, p.adj) %>% 
  arrange(Test, Group) %>%
  group_by(Test) %>% 
  #add positions
  mutate(hjustvar = 0) %>% 
  mutate(vjustvar = factor(Group, levels = unique(Group)) %>% 
           as.numeric + 0.5,
         vjustvar = vjustvar * 2) %>% 
  mutate(xpos = -Inf) %>% 
  mutate(ypos = Inf) %>% 
  mutate(annotateText = paste0("adj.p: ",round(p.adj, 3)))
```

```{r, fig.width=8}
biometrics %>%
  mutate(Timepoint = factor(Timepoint, levels = c("Before", "After"))) %>%
  filter(Test %in% (label.df %>% pull(Test) %>% unique)) %>%
  ggplot(aes (x = Timepoint, y = Value, group =  Group)) +
  geom_smooth(aes(col = Group, fill= Group), method = "glm", se = F) +
  geom_jitter(aes(col = Group)) +
  facet_wrap(~ Test, scales = "free", nrow = 1) +
  geom_text(data=label.df,aes(x = xpos,
                              y = ypos,
                              hjust = hjustvar,
                              vjust = vjustvar,
                              label = annotateText,
                              col = Group))
```

### Save plot elements

```{r}
save.figure.elements <- list(biometrics, label.df)
"figure.elements.rds" %>% 
  paste0(d.out, "/", .) %>% 
  saveRDS(save.figure.elements,.)
```


## All results
```{r, fig.width=10, fig.height=10}
biometrics %>%
  mutate(Timepoint = factor(Timepoint, levels = c("Before", "After"))) %>%
  ggplot(aes (x = Timepoint, y = Value, group =  Group)) +
  geom_smooth(aes(col = Group, fill= Group), method = "glm", se = F) +
  geom_jitter(aes(col = Group)) +
  facet_wrap(~ Test, scales = "free", nrow = 4) +
  geom_text(data=label.df,aes(x = xpos,
                              y = ypos, 
                              hjust = hjustvar, 
                              vjust = vjustvar,
                              label = annotateText, 
                              col = Group))
```


# Session information

```{r}
sessionInfo()
```
