---
title: "Run diversity analysis within a group"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
  group: "Control"
  stage: "intervention"
---


# Introduction
Run group-wise diversity analysis

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)

# Set output directory and data period
d.out <- params$d.out
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(vegan)
library(reshape2)
library(broom) # plotting residuas vs fit nicely
library(magrittr) #set colnames
library(spaa) #convert dist to list
library(plotrix) # calculates standard error
```

## Import data
```{r}
d.in <- "~/IKMB/projects/sport/data.clean"

meta <- "sampling.data.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read_tsv()
micro <-  "micro.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read.delim(stringsAsFactors = F)
micro.r <- "micro.r.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read.delim(stringsAsFactors = F)
alpha <- "alpha.tsv" %>%
  paste0(d.in, "/", .) %>% 
  read_tsv()
pca <- "~/IKMB/projects/sport/analysis/25.to.publish/6.diet/pca.rds" %>% 
  readRDS()
```

# Set working group and period

```{r}
group <- params$group
stage <- params$stage
```

# Filter data for group and period

## Filter metadata
```{r}
if(stage == "intervention"){
meta <- meta %>%
  filter(Group == group) %>% 
  filter(collection.date <= end.intervention)
}

if(stage == "after"){
meta <- meta %>%
  filter(Group == group) %>% 
  filter((collection.date > end.intervention -15) &
           (collection.date < end.intervention +15))
}
meta <- meta %>% 
  inner_join(pca, by = "Seq.ID")
```

## Filter microbiome

```{r}
micro <- micro %>%
  rownames_to_column("rownames") %>% # match microbiome data with meta
  left_join(data.frame(rownames = meta$Seq.ID),
            .,
            by = "rownames") %>%
  column_to_rownames("rownames")
micro <- micro[,colSums(micro)> 0]

micro.r <- micro.r %>%
  rownames_to_column("rownames") %>% # match microbiome data with meta
  left_join(data.frame(rownames = meta$Seq.ID),
            .,
            by = "rownames") %>%
  column_to_rownames("rownames")
micro.r <- micro.r[,colSums(micro.r)> 0]
```

## Filter alpha diversity

```{r}
alpha <- alpha %>%
  right_join(., meta,by = "Seq.ID")
```


# Visualize alpha diversity


```{r}
alpha %>%
  select(Seq.ID, Chao1, InvSimp) %>% 
  melt(variable.name = "alpha.div") %>%
  left_join(alpha,
            by = "Seq.ID") %>%
  ggplot(., aes(x = Days,
                     y = value,
                color = ParticipantID)) +
  geom_point() +
  geom_smooth(method = "lm", aes(group = alpha.div)) +
  theme(legend.position = "none") +
  facet_wrap(~ alpha.div, scales = "free") +
  labs(title = group)
```

## Stats

The aim is to see if there is any difference in alpha diversity due to time.


```{r}
#
do.lm <- function(df, variable){

f.f <- as.formula(paste("log(", variable, ")", "~", "ParticipantID + PC1 + Days"))

m.f <- lm(f.f, df)


res.fit.plot <- augment(m.f) %>% 
    ggplot(aes_string(x = ".fitted", y = ".resid")) +
    geom_point(alpha = 0.4, size = 5) + 
    geom_hline(yintercept=0,
               linetype="dashed", 
               color = "red",
               size = 1) +
    theme(legend.position="bottom")

print(res.fit.plot)

# Get p.value

sig <- anova(m.f) %>% 
  tidy() %>%
  mutate(index = variable,
         group = group,
         n = nrow(df)) 

# Calculate coefficient
model.coef <- m.f %>%
      coef %>%
      data.frame(coef = .) %>%
      rownames_to_column("term") %>%
      filter(term != "(Intercept)") %>%
      # get coefficient that matters
      filter(str_detect(term, "Days"))

 # Calculate coefficient intervals
model.coef.confi <- m.f %>%
  confint() %>%
  as.data.frame() %>%
  set_colnames(c("conf.low", "conf.high")) %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "Days")) 

r <-  sig %>%
    left_join(model.coef, by = "term") %>%
      left_join(model.coef.confi, by = "term")
  return(r)
}
```


### Perform tests
```{r}
alpha.stats <- NULL
alpha.stats <- bind_rows(alpha.stats,
                           do.lm(alpha,
                                     "Chao1"))
alpha.stats <- bind_rows(alpha.stats,
                           do.lm(alpha,
                                     "InvSimp"))
```

# Beta diversity


The aim is to see if there is any difference in beta diversity due to time.

##Making the plot

```{r}
mds <- micro.r %>%
  vegdist('bray') %>%
  metaMDS(.,autotransform = F)
```

## Visualize the NMDS plot


```{r, fig.height=12, fig.width=12}

mds$points %>%
  data.frame() %>%
  rownames_to_column("Seq.ID") %>%
  inner_join(alpha) %>%
  group_by(ParticipantID) %>%
  arrange(collection.date) %>%
  #plot
  ggplot(aes(x= MDS1, y = MDS2)) +
  #text
  geom_text(aes(colour = ParticipantID, group = ParticipantID, label =  Days), size = 3) +
  scale_shape_manual(values = c(16, 15)) +
  # add arrow
  geom_path(aes(group = ParticipantID, colour = ParticipantID),
            stat = "identity",
            arrow = arrow(length = unit(0.4, "cm"),
                          type = "open"),
            size = 1.1, alpha = 0.3) +
  #remove long legend
  guides(colour=FALSE) +
  #legend nicer, save space
  theme(legend.position="bottom") %>%
  labs(title = group)
```

```{r}
beta.stats <-  adonis2(micro.r ~ ParticipantID + PC1 + Days, data = alpha, by = "terms") %>%
  tidy() %>%
  mutate(group = group,
         method = "adonis.2")
```


## Save outputs

```{r}
paste0(d.out, "/", group, ".", stage , ".alpha.stats.tsv") %>%
  write_tsv(alpha.stats, .)


beta.stats <- beta.stats
paste0(d.out, "/", group, ".", stage , ".beta.stats.tsv") %>%
  write_tsv(beta.stats, .)
```


# Session information

```{r}
sessionInfo()
```
