---
title: "Summary of group-based analysis"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
  stage: "intervention"
---


# Introduction

I will gather group-based analyses for variations in microbiome along time.

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out

```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(reshape2)
```

## Import data
```{r}
d.in <- '/home/lsilva/IKMB/projects/sport/analysis/25.to.publish/7.microbiome'
d.in.data <- "~/IKMB/projects/sport/data.clean"
meta <- "sampling.data.tsv" %>%
  paste0(d.in.data, "/", .) %>% 
  read_tsv()
tax <- "tax.tsv"  %>%
  paste0(d.in.data, "/", .) %>% 
  read.delim(stringsAsFactors = F)
micro <-"micro.tsv" %>%
  paste0(d.in.data, "/", .) %>% 
  read.delim(stringsAsFactors = F)
```
# Set period

```{r}
stage <- params$stage
```


# Alpha diversity

## Get results
```{r}
alpha <- list.files(path = d.in,
                    pattern = paste0(stage,".alpha.stats.tsv")) %>%
  paste0(d.in, "/", .) %>%
  map(read_tsv) %>%
  do.call("bind_rows", .) %>%
  filter(term == "Days") %>% 
  arrange(p.value)
```


## Plot results
 Plot results (p-values) from the tests.
```{r}
alpha %>%
  filter(!is.na(p.value)) %>%
  mutate(Group = factor(group, levels = c("Control", "Endurance", "Strength"))) %>% 
  mutate(Direction = if_else(coef > 0 ,
                             "increase",
                             "decrease"),
         Coefficient = abs(coef)) %>%
  ggplot(., aes(x = index, y = 1 - p.value)) +
  geom_jitter(alpha = 0.5, aes(color = Group,
                               size = Coefficient,
                               shape = Direction),
              width = 0.1) +
  geom_hline(yintercept = 0.95, linetype="dashed") +
  labs(title = "Effect of time on alpha diversity",
       x = "Alpha diversity")
```

## View Significants

```{r}
alpha %>%
  filter(p.value < 0.05)
```


# Write alpha summary

```{r}
paste0("alpha.summary.",stage,".tsv") %>% 
  paste0(d.out, "/", .) %>% 
  write_tsv(alpha, .)
```

# Beta diversity

## Get results
```{r}
beta <- list.files(path = d.in,
                    pattern = paste0(stage, ".beta.stats.tsv")) %>%
  paste0(d.in, "/", .) %>%
  map(read_tsv) %>%
  do.call("bind_rows", .) %>%
  arrange(p.value)
```

## Plot results

Plot results (p-values) from the tests.
```{r, fig.width= 3}
beta %>%
  filter(term == "Days") %>% 
  filter(!is.na(p.value)) %>%
  mutate(group = factor(group, levels = c("Control", "Endurance", "Strength"))) %>% 
ggplot(aes(x = group, y = 1 - p.value)) +
  geom_point(alpha = 0.5, size = 3) +
  geom_hline(yintercept = 0.95, linetype="dashed") +
  ggpubr::rotate_x_text() +
  labs(title = "Effect of time\non beta-diversity",
       x = "Groups") +
  theme(legend.position = "none")
```

## View Significants

```{r}
beta %>%
  filter(p.value < 0.05)
```

# Write beta summary

```{r}
paste0("beta.summary.",stage,".tsv") %>% 
  paste0(d.out, "/", .) %>% 
  write_tsv(beta, .)
```

# ASVs that change along time

## Get results
```{r}
difab <- list.files(path = d.in,
                    pattern = paste0(stage, ".dseq.res.stats.tsv")) %>%
  paste0(d.in, "/", .) %>%
  map(read_tsv) %>%
  do.call("bind_rows", .) %>%
  arrange(padj)
```

## Plot results

Plot results (p-values) from the tests.
```{r}
difab %>% 
  mutate(Genus = gsub("_sensu_stricto", "", Genus)) %>% 
  mutate(group = factor(group, levels = c("Control", "Endurance", "Strength"))) %>% 
ggplot(aes(x = group, index, y = 1 - padj)) +
  geom_jitter(data = difab %>%  filter(padj >= 0.05) %>% 
                mutate(group = factor(group, levels = c("Control", "Endurance", "Strength"))),
              alpha = 0.5, width = 0.1, color = "black") +
  geom_jitter(data = difab %>%
                mutate(Direction =  if_else(log2FoldChange > 0 ,
                             "increase",
                             "decrease")) %>% 
                filter(padj < 0.05) %>% 
                mutate(Genus = gsub("_sensu_stricto", "", Genus)) %>% 
  mutate(group = factor(group, levels = c("Control", "Endurance", "Strength"))),
              alpha = 0.5, width = 0.1, aes(color = Genus,
                                            size = baseMean,
                                            shape = Direction)) +
  geom_hline(yintercept = 0.95, linetype="dashed") +
  ggpubr::rotate_x_text() +
  labs(title = "ASVs that vary along time",
       x = "Group")
```

## View Significants

```{r}
difab %>%
  filter(padj < 0.05)
```

# Write differential abudance summary

```{r}
paste0("diff.abundnace.summary.", stage, ".tsv") %>% 
  paste0(d.out, "/", .) %>% 
  write_tsv(difab, .)
```
# Session info
```{r}
sessionInfo()
```
