---
title: "Analyse diet"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
  group: "Endurance"
  stage: "after"
---


# Introduction
 I will analyze the change in diet patterns within each group.

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out
```

## Set working group
```{r}
group <- params$group
stage <- params$stage
rm(params)
```


## Load libraries for the session

```{r}
library(MASS)
library(ggplot2)
library(tidyverse)
library(vegan)
library(reshape2)
library(broom) # plotting residuas vs fit nicely
library(spaa) #convert dist to list
library(plotrix) # calculates standard error
library(factoextra)

```

## Import data
```{r}
d.in.pca <- "~/IKMB/projects/sport/analysis/25.to.publish/6.diet"

d.in.data <- "~/IKMB/projects/sport/data.clean" 

pca <- "pca.rds" %>% 
  paste0(d.in.pca, "/", .) %>% 
  readRDS()  

meta <- "sampling.data.tsv" %>% 
  paste0(d.in.data, "/", .) %>% 
  read_tsv()

dic <- "diet.dic.tsv" %>% 
  paste0(d.in.data, "/", .) %>% 
  read_tsv()

diet <- "diet.tsv" %>% 
  paste0(d.in.data, "/", .) %>% 
  read_tsv()
```

# Edit dictionary for a nice output
```{r}
dic <- dic %>% 
  separate(question, into = c("intake", "trash"), sep = " \\(", remove = F) %>% 
  dplyr::select(-trash)
```

# Add group information

```{r}
diet <- diet %>% 
  inner_join(meta,
             by = c("Seq.ID", "ParticipantID"))
```


# Filter for group and period
```{r}
if(stage == "intervention"){
diet <- diet %>%
  filter(Group == group) %>% 
  filter(Date <= end.intervention)
}

if(stage == "after"){
diet <- diet %>%
  filter(Group == group) %>% 
  filter((Date > end.intervention -15) &
           (Date < end.intervention +15))
}
```

# Remove points with missing diet info

```{r}
diet <- diet %>% 
  filter(diet %>% 
           select(dic %>%
                    pull(variable)) %>% 
           complete.cases())
```

# Organize dieta data

## Get metadata and diet information

```{r}
diet.meta <- diet
```

## Get data only with diet 

```{r}
diet <- diet %>% 
  select(dic %>%
           pull(variable) %>% 
           c("Seq.ID",.)) %>% 
  column_to_rownames("Seq.ID")
```

## View data
```{r}
diet %>%
  rownames_to_column("Seq.ID") %>% 
  melt(variable.name = "diet") %>%
  left_join(meta,
            by = "Seq.ID") %>%
  left_join(dic, by  = c("diet" = "variable")) %>% 
  rename("food" = "question") %>% 
  group_by(ParticipantID) %>%
  arrange(Days) %>% 
  ggplot(., aes(x = Days,
                     y = value,
                color = ParticipantID)) +
  geom_point() +
  geom_smooth(method = "lm", aes(group = intake)) +
  geom_path(aes(group = ParticipantID),
            stat = "identity",
            arrow = arrow(length = unit(0.15, "cm"),
                          type = "open"),
            size = 0.5, alpha = 0.3) +
  theme(legend.position = "none") +
  facet_wrap(~ intake, scales = "free") +
  labs(title = group)

```
# PCA

```{r}
pc.diet.meta <- pca %>% 
  right_join(diet.meta, by = "Seq.ID")

# Is it normally distributed?
hist(pc.diet.meta$PC1)

# Test
f <- as.formula(paste("PC1", "~", "ParticipantID + Days"))

m.f <- lm(f, pc.diet.meta)

res.fit.plot <- augment(m.f) %>% 
    ggplot(aes_string(x = ".fitted", y = ".resid")) +
    geom_point(alpha = 0.4, size = 5) + 
    geom_hline(yintercept=0,
               linetype="dashed", 
               color = "red",
               size = 1) +
    theme(legend.position="bottom")
res.fit.plot

sig <- anova(m.f) %>% 
  tidy() %>%
  mutate(index = "PC1",
         group = group,
         n = nrow(meta)) 

model.coef <- m.f %>%
      coef %>%
      data.frame(coef = .) %>%
      rownames_to_column("term") %>%
      filter(term != "(Intercept)") %>%
      # get coefficient that matters
      filter(str_detect(term, "Days"))

model.coef.confi <- m.f %>%
  confint() %>%
  as.data.frame() %>%
  magrittr::set_colnames(c("conf.low", "conf.high")) %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "Days")) 

sig <- sig %>%
    left_join(model.coef, by = "term") %>%
      left_join(model.coef.confi, by = "term")
sig
```

## Save PCA

```{r}
paste0(d.out, "/", group, ".intake.pca.", stage, ".tsv") %>%
  write_tsv(sig, .)

```

# Stats
The aim is to see if there is any difference in food intake along time. 
```{r}
do.lm <- function(df, variable){

f.f <- as.formula(paste("log(", variable, "+1)", "~", "ParticipantID + Days"))

m.f <- lm(f.f, df)


res.fit.plot <- augment(m.f) %>% 
    ggplot(aes_string(x = ".fitted", y = ".resid")) +
    geom_point(alpha = 0.4, size = 5) + 
    geom_hline(yintercept=0,
               linetype="dashed", 
               color = "red",
               size = 1) +
    theme(legend.position="bottom")

print(res.fit.plot)

# Get p.value

sig <- anova(m.f) %>% 
  tidy() %>%
  mutate(index = variable,
         group = group,
         n = nrow(df)) 

# Calculate coefficient
model.coef <- m.f %>%
      coef %>%
      data.frame(coef = .) %>%
      rownames_to_column("term") %>%
      filter(term != "(Intercept)") %>%
      # get coefficient that matters
      filter(str_detect(term, "Days"))

 # Calculate get coefficient confint
model.coef.confi <- m.f %>%
  confint() %>%
  as.data.frame() %>%
  magrittr::set_colnames(c("conf.low", "conf.high")) %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "Days")) 

r <-  sig %>%
    left_join(model.coef, by = "term") %>%
      left_join(model.coef.confi, by = "term")
  return(r)
}
```

## Perform tests

```{r}
stats.results <- NULL
for (i in colnames(diet)){
stats.results <- bind_rows(stats.results,
                           do.lm(diet.meta,
                                     i))}
```

## Add information

```{r}
stats.results <- stats.results %>% 
  left_join(dic, by = c("index" = "variable")) %>% 
  dplyr::select(intake, everything())
```


## Adjust p-value

```{r}
stats.results <- stats.results %>% 
  group_by(term) %>% 
  mutate(padj = p.adjust(p.value, "fdr")) %>% 
  dplyr::select(intake, p.value, padj, everything()) %>% 
  arrange(padj)
```

# Save

```{r}

paste0(d.out, "/", group, ".intake.stats.", stage, ".tsv") %>%
  write_tsv(stats.results, .)
```


# Session information

```{r}
sessionInfo()
```
