---
title: "Summary of group-based analysis"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
  stage: "intervention"
---


# Introduction

I will gather group-based analyses for variations in food intake along time.
# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}



# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out
stage <- params$stage
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(reshape2)

```

## Import data
```{r}
d.in <- '/home/lsilva/IKMB/projects/sport/analysis/25.to.publish/6.diet'

d.in.data <- "~/IKMB/projects/sport/data.clean" 

meta <- "sampling.data.tsv" %>% 
  paste0(d.in.data, "/", .) %>% 
  read_tsv()

dic <- "diet.dic.tsv" %>% 
  paste0(d.in.data, "/", .) %>% 
  read_tsv()

diet <- "diet.tsv" %>% 
  paste0(d.in.data, "/", .) %>% 
  read_tsv()
```

# Add metadata to diet data

```{r}
diet.meta <- diet %>% 
  inner_join(meta,
             by = c("Seq.ID", "ParticipantID"))
```



# Individual dietary components

## Get results
```{r, echo = T, results = 'hide', message=F}
intake <- list.files(path = d.in,
                     pattern = paste0("intake.stats.", stage, ".tsv")) %>%
  paste0(d.in, "/", .) %>%
  map(read_tsv) %>%
  do.call("bind_rows", .) %>%
  arrange(p.value)
```

## Include only the results for Days
```{r}
intake <- intake %>%
  filter(term == "Days") 
```


## Plot results for figure
```{r, fig.height= 8, fig.width=8}
g1 <- intake %>%
  filter(term == "Days") %>% 
  filter(group %in% c("Control", "Endurance", "Strength") ) %>% 
  filter(!is.na(p.value)) %>%
  mutate(Group = factor(group, levels = c("Control", "Endurance", "Strength"))) %>% 
  mutate(Direction = if_else(coef > 0 ,
                             "increase",
                             "decrease"),
         Direction = if_else(is.na(Direction),
                             "absent",
                             Direction),
         Direction = factor(Direction, levels = c("decrease",
                                                  "increase",
                                                  "absent")),
         Coefficient = abs(coef)) %>%
  mutate(intake = str_wrap(intake, 20)) %>% 
  
  ggplot(., aes(x = intake, y = 1 - padj)) +
  geom_point(alpha = 0.8, size =0.1) +
  geom_point(alpha = 0.8, aes(color = Group,
                              size = Coefficient,
                              shape = Direction)) +
  geom_hline(yintercept = 0.95, linetype="dashed") +
  labs(x = "Dietary component", y = " 1 - adjusted P")+
  ggpubr::rotate_x_text() 
g1
```

## View Significants

```{r}
intake %>%
  filter(term == "Days") %>% 
  select(intake, group, coef, everything()) %>% 
  filter(padj < 0.05) %>% 
  arrange(group)
```

# Plot the data

## make label
```{r}
label.df <- intake %>%
  filter(term == "Days") %>% 
  select(intake, group, coef, everything()) %>% 
  filter(padj < 0.05) %>% 
  arrange(group) %>% 
  rename("Group" = "group") %>% 
  mutate(label = "*") %>% 
  select(Group, label, padj) %>% 
  arrange(Group) %>%
  #add positions
  mutate(hjustvar = 0) %>% 
  mutate(vjustvar = 0.5,
         vjustvar = vjustvar * 2) %>% 
  mutate(xpos = -Inf) %>% 
  mutate(ypos = Inf) %>% 
  mutate(annotateText = paste0(" ",round(padj, 3), " (i)"))


```


```{r}
g2 <- diet.meta %>% 
  select(Mcc.p.2e2ssoh, ParticipantID, Group, Date, Days) %>% 
  ggplot(aes(x = Days, y = Mcc.p.2e2ssoh)) +
  facet_wrap(~ Group) + #Group ~ ParticipantID) +
  geom_vline(xintercept = 0 + 20, 
             color = "black",
             size=0.3) +
  geom_vline(xintercept = 42 + 20, 
             color = "black",
             size=0.5) +
  ylim(low = min(diet.meta$Mcc.p.2e2ssoh), high = max(diet.meta$Mcc.p.2e2ssoh)) +
  geom_point(alpha = 0.5, color = "grey70") +
  geom_path(color = "grey70", aes(group = ParticipantID),
            stat = "identity",
            arrow = arrow(length = unit(0.15, "cm"),
                          type = "open"),
            size = 0.5, alpha = 0.2) +
  # Fit statistical model for the intervention
  geom_smooth(data = diet.meta %>%
                filter(Date <= end.intervention),
              method = "lm", aes(color = Group, fill=Group),
              formula = lm(Mcc.p.2e2ssoh ~ ParticipantID + Days,
                           data = diet.meta %>%
                             filter(Date <= end.intervention))$call %>% str()) +
  
  # Fit statistical model after intervention
  geom_smooth(data = diet.meta %>%
                filter((Date > end.intervention -15) &
                         (Date < end.intervention +15)),
              method = "lm", aes(color = Group, fill=Group), linetype = "dashed",
              formula = lm(Mcc.p.2e2ssoh ~ ParticipantID + Days, 
                           data = diet.meta %>%
                             filter((Date > end.intervention -15) &
                                      (Date < end.intervention +15)))$call %>% 
                str()) +
  labs(x = "Days", y = "Meat cold cuts\n(1 portion = 20g)") +
  geom_text(data=label.df,aes(x = xpos,
                              y = ypos,
                              hjust = hjustvar,
                              vjust = vjustvar,
                              label = annotateText, 
                              col = Group)) +
  theme(legend.position = "none")
g2
```



# Combine PC1 results

```{r}
intake <- list.files(path = d.in,
                     pattern = paste0("intake.pca.", stage, ".tsv")) %>%
  paste0(d.in, "/", .) %>%
  map(read_tsv) %>%
  do.call("bind_rows", .) %>% 
  filter(group != "Extreme") %>% 
  mutate(term = case_when(term == "ParticipantID" ~ "Participant ID",
                          term == "Days" ~ "Days")) %>% 
  select(group, term, everything(), -index )
colnames(intake) <- str_to_sentence(colnames(intake))

colnames(intake) <- str_replace(colnames(intake), "\\.", " ") 

colnames(intake) <- str_replace(colnames(intake), "Pct explained", "Explained (%)") 

kable(intake, digits = 3)

```

# Plot coefs

```{r}
label.df <- intake %>%
  filter(Term == "Days") %>% 
  filter(`P value` < 0.05) %>% 
  arrange(Group) %>% 
  mutate(label = "*") %>% 
  select(Group, label,`P value`) %>% 
  arrange(Group) %>%
  #add positions
  mutate(vjustvar = 1) %>% 
  mutate(hjustvar = -1) %>% 
  mutate(ypos = -Inf) %>% 
  mutate(xpos = Inf) %>% 
  mutate(annotateText = paste0(" ",round(`P value`, 3), " (i)"))



g3 <- intake %>% 
  filter(Term == "Days") %>% 
  ggplot(aes(x = Group, y = Coef,color = Group)) +
  geom_point() +
  coord_flip() +
  geom_errorbar(aes(ymin=`Conf low`, ymax=`Conf high`), width=.2,
                position=position_dodge(.9)) +
  geom_hline(yintercept = 0) +
  labs(x = "Coefficient") +
  geom_text(data=label.df,aes(x=xpos,y=ypos,hjust=hjustvar,vjust=vjustvar,label=annotateText, col = Group)) +
  theme(legend.position = "none") +
  labs(y = "PC1\nβ coefficient", x = "Group")
g3
```

# Save plots

```{r}
paste0("plots.", stage, ".rds") %>% 
  paste0(d.out, "/", .) %>% 
  saveRDS(list(g1,g2,g3), .)

```


# Session information

```{r}
sessionInfo()
```
