---
title: "Visualize and test sleep data"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction

I will visualize and test sleeping data.

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)
# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(reshape2)
library(RColorBrewer)
library(ggnewscale)
library(broom) # plotting residuas vs fit nicely
library(magrittr) #set colnames
library(spaa) #convert dist to list
library(plotrix) # calculates standard error
```

## Import data
```{r}
meta <- "~/IKMB/projects/sport/data.clean/sampling.data.tsv" %>% 
  read_tsv()
sleep <- "~/IKMB/projects/sport/data.clean/sleep.tsv" %>% 
  read_tsv()
```

## Add Groups and intervention dates

```{r}
sleep <- sleep %>% 
  inner_join(meta %>%
               select(ParticipantID, start.intervention, end.intervention, Group) %>% 
               unique(),
             by = "ParticipantID") %>% 
  mutate(min = seconds /60) %>% 
  mutate(hour = min/60)
```

## Convert date to days after intervention

```{r}
sleep <- sleep %>% 
  mutate(days.after = Date - start.intervention) %>% 
  mutate(days.after.int = as.numeric(days.after) + 20)
```

# Data filtering
## Remove participants with few recorded days
cut-off is : 10

```{r}
sleep <- sleep %>%
  group_by(ParticipantID) %>% 
  mutate(n = n()) %>% 
  filter(n > 10)
```

## Remove data that does not match the study length
```{r}
sleep <- sleep %>% 
  filter(days.after.int < 70)
```

# Plot average sleep per day

```{r}
g1 <- sleep %>%
  group_by( ParticipantID, Group) %>% 
  summarise(avg.hour = mean(hour), sd = sd(hour)) %>% 
  ggplot(aes(x = Group, y = avg.hour)) +
  geom_boxplot(coef=6) +
  geom_jitter(aes(color=Group)) +
  labs(y = "Average hours") +
  ggpubr::rotate_x_text()
g1
```
# Save sleep

```{r}
"sleep.rds" %>% 
  paste0(d.out, "/",.) %>% 
  saveRDS(sleep, .)
```


#Print summary

```{r}
sleep %>%
  group_by( ParticipantID, Group) %>% 
  summarise(avg.hour = mean(hour), sd = sd(hour)) %>% 
  group_by(Group) %>% 
  summarise(sd = sd(avg.hour),median = median(avg.hour), avg = mean(avg.hour)) %>% 
  arrange(median, sd, avg) %>% 
  mutate(view = paste0(round(avg,2), "±", round(sd,2)))
```


# Analyse sleep along days

##  Function

```{r}
do.lm <- function(df, variable){
  
  df <- df %>% 
    filter(Group == variable)
  
  f.f <- as.formula("hour ~ ParticipantID + days.after.int")
  
  m.f <- lm(f.f, df)
  
  
  res.fit.plot <- augment(m.f) %>% 
    ggplot(aes_string(x = ".fitted", y = ".resid")) +
    geom_point(alpha = 0.4, size = 5) + 
    geom_hline(yintercept=0,
               linetype="dashed", 
               color = "red",
               size = 1) +
    theme(legend.position="bottom")
  
  print(res.fit.plot)
  
  # Get p.value
  
  sig <- anova(m.f) %>% 
    tidy() %>%
    mutate(index = variable,
           n = nrow(df)) 
  
  # Calculate coefficient
  model.coef <- m.f %>%
    coef %>%
    data.frame(coef = .) %>%
    rownames_to_column("term") %>%
    filter(term != "(Intercept)") %>%
    # get coefficient that matters
    filter(str_detect(term, "days.after.in"))
  
  # Calculate get coefficient confint
  model.coef.confi <- m.f %>%
    confint() %>%
    as.data.frame() %>%
    set_colnames(c("conf.low", "conf.high")) %>%
    rownames_to_column("term") %>%
    filter(str_detect(term, "days.after.int")) 
  
  r <-  sig %>%
    left_join(model.coef, by = "term") %>%
    left_join(model.coef.confi, by = "term")
  return(r)
}

```

## Perform test

# Intervention
```{r}
stats.results.int <- NULL
stats.results.int <- bind_rows(stats.results.int,
                               do.lm(sleep %>%
                                       filter(Date <= end.intervention),
                                     "Strength"))
stats.results.int <- bind_rows(stats.results.int,
                               do.lm(sleep %>%
                                       filter(Date <= end.intervention),
                                     "Endurance"))
stats.results.int <- bind_rows(stats.results.int,
                               do.lm(sleep %>%
                                       filter(Date <= end.intervention),
                                     "Control"))
```


# After intervention

```{r}
stats.results.after <- NULL
stats.results.after <- bind_rows(stats.results.after,
                                 do.lm(sleep %>%
                                         filter((Date > end.intervention -15) &
                                                  (Date < end.intervention +15)),
                                       "Strength"))
stats.results.after <- bind_rows(stats.results.after,
                                 do.lm(sleep %>%
                                         filter((Date > end.intervention -15) &
                                                  (Date < end.intervention +15)),
                                       "Endurance"))
stats.results.after <- bind_rows(stats.results.after,
                                 do.lm(sleep %>%
                                         filter((Date > end.intervention -15) &
                                                  (Date < end.intervention +15)),
                                       "Control"))
```


# Combine tests

```{r}
stats <- bind_rows(stats.results.int %>%
                     mutate(type = "intervention"),
                   stats.results.after %>%
                     mutate(type = "after"))
```

# Write results

```{r}
"stats.results.tsv" %>% 
  paste0(d.out, "/", .) %>% 
  write_tsv(stats, .)
```

# Plot

### Plot 

####  Make label
```{r}
label.df <- stats %>%
  filter(p.value < 0.05,
         term == "days.after.int") %>% 
  rename("Group" = "index") %>% 
  mutate(label = "*") %>% 
  select(Group, label, p.value, type) %>% 
  arrange(Group) %>%
  #add positions
  mutate(hjustvar = 0) %>% 
  mutate(vjustvar = 0.5,
         vjustvar = vjustvar * 2) %>% 
  mutate(xpos = -Inf) %>% 
  mutate(ypos = Inf) %>% 
  mutate(annotateText = paste0(" ",round(p.value, 3), " ", str_sub(type, 1, 3)), ".")

```


```{r}
g2 <- sleep %>% 
  ggplot(aes(x = days.after.int, y = hour)) +
  facet_wrap(~ Group) + #Group ~ ParticipantID) +
  geom_vline(xintercept = 0 + 20, 
             color = "black",
             size=0.3) +
  geom_vline(xintercept = 42 + 20, 
             color = "black",
             size=0.5) +
  ylim(low = min(sleep$hour), high = max(sleep$hour)) +
  geom_point(alpha = 0.5, color = "grey70") +
  geom_path(color = "grey70", aes(group = ParticipantID),
            stat = "identity",
            arrow = arrow(length = unit(0.15, "cm"),
                          type = "open"),
            size = 0.5, alpha = 0.2) +
  # Fit statistical model for the intervention
  geom_smooth(data =sleep %>%
                filter(Date <= end.intervention),
              method = "lm", aes(color = Group, fill=Group),
              formula = lm(hour ~ ParticipantID + days.after.int, data = sleep %>%
                             filter(Date <= end.intervention))$call %>% str()) +
  
  # Fit statistical model after intervention
  geom_smooth(data = sleep %>%
                filter((Date > end.intervention -15) &
                         (Date < end.intervention +15)),
              method = "lm", aes(color = Group, fill=Group), linetype = "dashed",
              formula = lm(hour ~ ParticipantID + days.after.int, data = sleep %>%
                             filter((Date > end.intervention -15) &
                                      (Date < end.intervention +15)))$call %>% str()) +
  labs(x = "Days", y = "Hours")+
  geom_text(data=label.df,aes(x=xpos,y=ypos,hjust=hjustvar,vjust=vjustvar,label=annotateText, col = Group)) +
  theme(legend.position = "none")
g2
```



# Session information

```{r}
sessionInfo()
```
